Adaptive Waveform Inversion for Transmitted Wavefields

### 2024-02-15 repro/papers/GEM22/submitR2/paper.tex
cleaned up

The present paper concerns {\em surface source extension}: physical
sources are presumed to be concentrated at points in space, whereas
their extended counterparts are permitted to spread energy over a {\em
  source surface}. Similarly, receivers are located on a {\em receiver
  surface}. Assuming that the presumed point source is at location
$\bx_s$, a simple (though certainluy not the only) choice for the
penalty operator $A$ is then multiplication by the distance
$|\bx-\bx_s|$ to the physical source location:
\begin{equation}
  \label{eqn:penop}
  (A\bff)(\bx,t) = |\bx-\bx_s|\bff(\bx,t)
\end{equation}
I shall use this choice of penalty operator whenever a specific choice
is necessary in the development of the theory below.

>>>>

The present paper concerns {\em surface source extension}: physical
sources are presumed to be concentrated at points in space, whereas
their extended counterparts are permitted to spread energy over a {\em
  source surface}. Similarly, receivers are located on a {\em receiver
  surface}. Assuming that the presumed point source is at location
$\bx_s$, a simple (though certainly not the only) choice for the
penalty operator $A$ is then multiplication by the distance
$|\bx-\bx_s|$ to the physical source location:
\begin{equation}
  \label{eqn:penop}
  (A\bff)(\bx,t) = |\bx-\bx_s|\bff(\bx,t)
\end{equation}
I shall use this choice of penalty operator whenever a specific choice
is necessary in the development of the theory below.

### 2024-02-15 repro/papers/awi/paper.tex
cleaned up

\begin{abstract}
Under some restricted circumstances, Adaptive Waveform Inversion
applied to transmitted wave data yields estimates of index of
refraction (or wave velocity) similar to those obtained by travel time inversion.
\end{abstract}

\section{Introduction}
Full waveform inversion (FWI) refers to the inference of mechanical
parameter distribution within a material object  (the earth, a human
body, a manufactured artifact,...) from the measurement of wave motion
at its surface. Over the last several decades, this task has come to
be carried out by minimizing the difference between computationa
prediction of wave measurement versus actual observed
measurements. The mechanical parameters to be inferred appear as
coefficients in systems of partial differential equaitons governing
the wave motion, and discrete representations of these parameters form
the unknown vectors in the optimization problems thus generated.

@@

Various modifications of straighforward best-fit FWI have been
suggested to overcome this ``cycle-skipping'' pathology. One of these
is Adaptive Waveform Inversion \cite[]{Warner:16} (AWI). This
technique has been successfully applied to industry-scale seismic data
\cite[]{GuaschWarnerRavaut:GEO19,Warneretal:SEG21} as well as to
ultrasonic imaging of the the human brain
\cite[]{Guaschetal:NPJDM20}.

While heuristic justifications for AWI (and
many details of implementation) are discussed in these and other
works, a clear explanation for its effectiveness is 
so far missing from the literature. In this short note, I show that
for active sensing of acoustic materials with unique arrival times of
transmitted waves
(that is unique rays of geometric asymptotics between sources and
receivers), the AWI objective function is an approximation to the
mean-square time-of-travel misfit. Thus AWI and travel time fitting
lead to similar material parameter estimates. Since the latter is
widely believed to be a near-optimal method for wave velocity
estimation, this demonstration explains roughly ``why it works''.

The reader will
note that a normalization component of the AWI objective function
definition is very
important in reaching this conclusion. The connection between
objective normalization and the close relation to travel time
inversion is completely independent from 
the reasons for normalizing the objective advanced in the
literature \cite[]{Warner:16,LiAlkhalifah:21}.

@@

Acoustic inversion, for present purposes, means: given recording time interval
$[0,T]$, source and receiver locations $(\bx_s,\bx_r)$, source wavelet
$w(t;\bx_s)$ and data traces $d(\bx_r,\cdot;\bx_s)$, find a model $m$
so that $F[m]w \approx d$. The
simplest version of FWI concretizes this task by asking for a model
$m$ minimizing the mean square error
\begin{equation}
  \label{eqn:fwi}
  J_{\rm FWI}[m;d] =\frac{1}{2} \sum_{{\bf x}_r,{\bf x}_s} \int\,dt\, |(F[m]w)({\bf x}_r,t;{\bf x}_s) - d({\bf x}_r,t;{\bf x}_s)|^2
\end{equation}
It is convenient to introduce the $L^2$ norm $\| \cdot \|$ as
shorthand for the right-hand side of the definition \ref{eqn:fwi}:
\[
  J_{\rm FWI}[m;d]= \frac{1}{2}\|F[m]w-d\|^2.
\]
In the sequel I will use the $L^2$ norm notation to stand for
mean-square over whatever variables make sense.

Note that the wavelet $w$ is assumed known in this formulation of the
FWI problem. In fact, even if actual acoustic sources are
well-approximated by isotropic point radiators, their wave forms are
usually poorly determined and may reasonably be
included as  parameters to be estimated by inversion of the
data. However for the present discussion we will regard $w$ as known.

>>>>

\begin{abstract}
Adaptive Waveform Inversion
applied to transient transmitted wave data yields estimates of index of
refraction (or wave velocity) similar to those obtained by travel time
inversion, provided that the data contain a single smooth wavefront.
\end{abstract}

\section{Introduction}
Full waveform inversion (FWI) refers to the inference of mechanical
parameter distribution within a material object  (the earth, a human
body, a manufactured artifact,...) from the measurement of wave motion
at its surface. Over the last several decades, this task has come to
be carried out by minimizing the difference between computationa
prediction of wave measurement versus actual observed
measurements. The mechanical parameters to be inferred appear as
coefficients in systems of partial differential equations governing
the wave motion, and discrete representations of these parameters form
the unknown vectors in the optimization problems thus generated.

@@

Various modifications of straighforward best-fit FWI have been
suggested to overcome this ``cycle-skipping'' pathology. One of these
is Adaptive Waveform Inversion \cite[]{Warner:16} (AWI). This
technique has been successfully applied to industry-scale seismic data
\cite[]{GuaschWarnerRavaut:GEO19,Warneretal:SEG21} and shows promise for
ultrasonic imaging of the the human brain
\cite[]{Guaschetal:NPJDM20}.

While heuristic justifications for AWI (and
many details of implementation) are discussed in these and other
works, a clear explanation for its effectiveness is 
so far missing from the literature. In this short note, I show that
for active sensing of acoustic materials with unique arrival times of
transmitted waves
(that is, unique rays of geometric asymptotics between sources and
receivers), the AWI objective function is an approximation to the
mean-square time-of-travel misfit. Thus AWI and travel time fitting
lead to similar material parameter estimates. Since the latter is
widely believed to be a near-optimal method for wave velocity
estimation, this demonstration explains roughly ``why AWI works'', at
least within this restricted scope.

The reader will
note that a normalization component of the AWI objective function
definition is very
important in reaching this conclusion. The connection between
objective normalization and the close relation to travel time
inversion has nothing to do with
the reasons for normalizing the objective advanced in the
literature \cite[]{Warner:16,LiAlkhalifah:21}.

@@

Acoustic inversion, for present purposes, means: given source and receiver locations $(\bx_s,\bx_r)$, source wavelet
$w(t;\bx_s)$ and data traces $d(\bx_r,\cdot;\bx_s)$, find a model $m$
so that $F[m]w \approx d$. The
simplest version of FWI concretizes this task by asking for a model
$m$ minimizing the mean square error
\begin{equation}
  \label{eqn:fwi}
  J_{\rm FWI}[m;d] =\frac{1}{2} \sum_{{\bf x}_r,{\bf x}_s} \int\,dt\, |(F[m]w)({\bf x}_r,t;{\bf x}_s) - d({\bf x}_r,t;{\bf x}_s)|^2
\end{equation}
It is convenient to introduce the $L^2$ norm $\| \cdot \|$ as
shorthand for the right-hand side of the definition \ref{eqn:fwi}:
\[
  J_{\rm FWI}[m;d]= \frac{1}{2}\|F[m]w-d\|^2.
\]
In the sequel I will use the $L^2$ norm notation to stand for
mean-square over whatever variables make sense.

Note that in actual data acquisition, recording occurs over finite
time intervals, so the integrals appearing in the definition
\ref{eqn:fwi} have finite upper and lower limits, as opposed to the
$\pm \infty$ limits implicit in the notation. It will be assumed that
wavefield energy is causal (zero before an onset time) and decays
rapidly enough as $t \rightarrow \infty$ that the difference between
finite and infinite range of integration can be ignored throughout
this discussion.

The wavelet $w$ is assumed known in this formulation of the
FWI problem. In fact, even if actual acoustic sources are
well-approximated by isotropic point radiators, their wave forms are
usually poorly determined and may reasonably be
included as  parameters to be estimated by inversion of the
data. However for the present discussion we will regard $w$ as known.
